---
title: "Hierarchical model"
author: "Rabeez Riaz"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cmdstanr)
register_knitr_engine()
```

## Motivation

To model discrete heterogeneity we can use hierarchical (multi-level) models to effectively incorporate information from multiple groups. We will use a Bayesian hierarchical model. This approach has 3 main advantages:
1. The Bayesian modelling approach allows for setting of priors which can incorporate domain knowledge
2. Due to the 'pooling' effect the estimated parameters are more stable
3. If a certain group has a small sample size, we can still estimate parameters for it reasonably well assuming the priors are not completely horrible because we can incorporate information from the other groups with larger samples


https://www.youtube.com/watch?v=dNZQrcAjgXQ

## Dataset

premier league 1920 matches
20 teams
328 matches
home/away (home advantage)


```{r}
df <- read_csv(
  "https://raw.githubusercontent.com/MaggieLieu/STAN_tutorials/master/Hierarchical/premiereleague.csv",
  col_names = c("home_team", "home_score", "away_score", "away_team")
)
df

```



## Model definition
number of goals ~ poisson (count)
where rate parameter of poisson (theta) is the expected score

for a match between TEAM1 (home) and TEAM2 (away) the expected scores are:
theta_TEAM1 = exp(home_advantage_factor + attack_ability_TEAM1 - defense_ability_TEAM2)
theta_TEAM2 = exp(attack_ability_TEAM2 - defense_ability_TEAM1)
^ take exp to make positive


```{stan, output.var="model"}
data {
  int<lower=0> num_teams;
  int<lower=0> num_games;
  int<lower=0> home_team_idx[num_games];
  int<lower=0> away_team_idx[num_games];
  int<lower=0> home_team_score[num_games];
  int<lower=0> away_team_score[num_games];
}

parameters {
  real home_adv;
  vector[num_teams] attack_ability;
  vector[num_teams] defense_ability;
  
  // hyper parameters
  real mu_attack;
  real<lower=o> tau_attack;
  real mu_defense;
  real<lower=o> tau_defense;
}

transformed parameters {
  vector[num_games] theta1;
  vector[num_games] theta2;
  
  theta1 = exp(home_adv + attack_ability[home_team_idx] - defense_ability[away_team_idx]);
  theta2 = exp(attack_ability[away_team_idx] - defense_ability[home_team_idx]);
}

model {
  // hyper priors
  mu_attack ~ normal(0, 0.1);
  tau_attack ~ normal(0, 1);
  mu_defense ~ normal(0, 0.1);
  tau_defense ~ normal(0, 1);
  
  // priors
  attack_ability ~ normal(mu_attack, tau_attack);
  defense_ability ~ normal(mu_defense, tau_defense);
  home_adv ~ normal(0, 0.001);
  
  // likelihood
  home_team_score ~ poisson(theta1);
  away_team_score ~ poisson(theta2);
}
```



## Model fitting

```{r}
model$sample()
```


## Findings


```{r}
names(model)
```


## Takeaways

